---
title: "ICU Report"
format: gfm
execute: 
  eval: true
  echo: false
  warning: false
  message: false
editor: visual
---

Report last run: `r Sys.time()`

## Introduction
This report is based on the most recent physiological data about the ICU patients being tracked.

```{r, echo=FALSE}
## setup libraries and data
library(tidyverse)
library(data.table)
library(lubridate)
library(googledrive)
library(readr)
source("sepsis_monitor_functions.R")
## Calling drive_deauth() prevents R from trying to authenticate via a browser
## This is needed to make the GitHub Action work
drive_deauth()
file_link <- "https://drive.google.com/file/d/1tumqraj4-TYNViaG6abUBl8Tx4pMN-wU/view?usp=share_link"
## All data up until now
new_data <- updateData(file_link)
## Include only most recent data
most_recent_data <- new_data %>%
  group_by(PatientID) %>%
  filter(obsTime == max(obsTime))
```


## A table listing the patients who currently have sepsis (if any), along with their most recent heart rate, temperature, and respiratory rate

```{r}
library(gt)
## check if there are any patients with sepsis and print message if not
if (sum(new_data$SepsisLabel == 1) == 0) {
  print("There are no patients with sepsis at the time of this report")
} else {
## create a table with patientID, HR, Temp, and Resp for patients with sepsis
# filter patients with sepsis and select columns
  sepsis_table_data <- new_data %>%
    filter(SepsisLabel == 1) %>%
    select(PatientID, HR, Temp, Resp)
  
# create gt table
  gt(sepsis_table_data) %>%
    tab_header("ICU Patients with Sepsis") %>%
      cols_label(PatientID = "Patient ID", 
             HR = "Heart Rate", 
             Temp = "Temperature", 
             Resp = "Respiratory Rate")
}
```



## Plots for all patients who have sepsis (if any) that show the complete history of their heart rate, temperature, and respiratory rate during their time in the ICU.

```{r}
## check if there are any patients with sepsis and print message if not
if (sum(new_data$SepsisLabel == 1) == 0) {
  print("There are no patients with sepsis at the time of this report")
} else {
## create a ggplot with HR, Temp, and Resp history for patients with sepsis
## filter sepsis pts
  sepsis_plot_data <- new_data %>%
    filter(SepsisLabel == 1) %>%
    group_by(PatientID)
    
## Heart Rate
    ggplot(sepsis_plot_data, aes(x = ICULOS, y = HR, color = PatientID)) +
      geom_line() +
      labs(title = "Heart Rate History", x = "Time in ICU", y = "Heart Rate")
    
## Temperature
    ggplot(sepsis_plot_data, aes(x = ICULOS, y = Temp, color = PatientID)) +
      geom_line() +
      labs(title = "Heart Rate History", x = "Time in ICU", y = "Temperature")
    
## Respiratory Rate
        ggplot(sepsis_plot_data, aes(x = ICULOS, y = Resp, color = PatientID)) +
      geom_line() +
      labs(title = "Heart Rate History", x = "Time in ICU", y = "Respiratory Rate")
  
}
```


## A table showing the change in heart rate, temperature, and respiratory rate between the last two measurements for all patients
```{r}
## Pull last two measurements for each patient in the new data
last_two_HR <- new_data %>%
  select(PatientID, HR) %>%
  group_by(PatientID) %>%
  filter(!is.na(HR)) %>%
  slice_tail(n = 2) %>%
  summarize(HR_change = diff(HR, na.rm = TRUE))
last_two_temp <- new_data %>%
  select(PatientID, Temp) %>%
  group_by(PatientID) %>%
  filter(!is.na(Temp)) %>%
  slice_tail(n = 2) %>%
  summarize(Temp_change = diff(Temp, na.rm = TRUE))
last_two_Resp <- new_data %>%
  select(PatientID, Resp) %>%
  group_by(PatientID) %>%
  filter(!is.na(Resp)) %>%
  slice_tail(n = 2) %>%
  summarize(Resp_change = diff(Resp, na.rm = TRUE))
##join data by PatientID
vitals_change <- full_join(last_two_HR, last_two_temp, by = "PatientID") %>%
  full_join(last_two_Resp, by = "PatientID")
##Create Table with change in values
gt(vitals_change) %>%
  tab_header("ICU Patients with Sepsis") %>%
  cols_label(PatientID = "Patient ID", 
             HR_change = "Heart Rate Change", 
             Temp_change = "Temperature Change", 
             Resp_change = "Respiratory Rate Change")
```